What We Eval in the Shadows

### 2021-04-05 analysis/normalized.Rmd
Add examples for normalized expressions
Provenance of eval arguments

Minimized:
```{r}
widespread_minimized <- E_cano %>%
  group_by(minimized) %>%
  summarize(n_call_sites = n_distinct(src_ref), 
            proportion = ratio(n_call_sites, nb_call_sites), 
            n_packages = n_distinct(ev_package),
            n_operations = round(mean(n_op, na.rm=TRUE), 2), 
            percent_envir = ratio(n_distinct(cur_data()$src_ref[!is.na(cur_data()$envir)]),  n_call_sites), 
            percent_envir_calls = ratio(sum(!is.na(envir)), n()), 
            env_diversity = n_distinct(envir), 
            percent_envir_non_parentframe = ratio(n_distinct(cur_data()$src_ref[!is.na(cur_data()$envir) & cur_data()$envir != "parent.frame()"]),  n_call_sites)) %>% 
  arrange(desc(n_call_sites))

@@

>>>>

Minimized:
```{r}
widespread_minimized <- E_cano %>%
  group_by(minimized) %>%
  summarize(n_call_sites = n_distinct(src_ref), 
            proportion = ratio(n_call_sites, nb_call_sites), 
            n_packages = n_distinct(ev_package),
            n_operations = round(mean(n_op, na.rm=TRUE), 2), 
            med_operations = median(n_op, na.rm = TRUE),
            percent_envir = ratio(n_distinct(cur_data()$src_ref[!is.na(cur_data()$envir)]),  n_call_sites), 
            percent_envir_calls = ratio(sum(!is.na(envir)), n()), 
            env_diversity = n_distinct(envir), 
            percent_envir_non_parentframe = ratio(n_distinct(cur_data()$src_ref[!is.na(cur_data()$envir) & cur_data()$envir != "parent.frame()"]),  n_call_sites)) %>% 
  arrange(desc(n_call_sites))

@@

# Examples of minimized expression

```{r}
examples_minimized <- E_cano %>% group_by(minimized) %>% filter(!is.na(expr_resolved) & str_length(expr_resolved) < 50) %>% 
  summarise(example = nth(unique(expr_resolved), 100))

examples_minimized %>% datatable()
```

### 2021-04-05 analysis/provenance.Rmd
Provenance of eval arguments

## Expression tyoes

@@

parsed_exprs <- E_prov %>% filter(!is.na(parsed))

@@

str2lang_exprs <- language %>% filter(!is.na(parsed))

@@

>>>>

## Expression types

@@

parsed_exprs <- expressions %>% filter(!is.na(parsed))

@@

str2lang_exprs <- languages %>% filter(!is.na(parsed))

@@

## Sites wwith less signficant provenance

```{r}
unknown_prov <- E_prov %>% anti_join(expressions) %>% anti_join(languages) %>% anti_join(bytecode_exprs)

nb_unknown_prov_sites <- unknown_prov %>% pull(src_ref) %>% n_distinct()

overview_table(
    r("nb unknown prov sites", nb_unknown_prov_sites),
    r("unknown prov site percent", ratio(nb_unknown_prov_sites, nb_call_sites))
)
```

### 2021-04-05 paper/main.tex
Provenance of eval arguments

@@

\mypara{Source.} % rename to provenance

There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

We observe 5.2\% cases where \eval directly evaluates the output of
\c{substitute}, 0.7\% cases where output of \c{parse} is read directly
and only 704 cases for \c{expression}. Most expressions consumed by
\eval are generated by other functions.

\c{eval(parse(...))} can be used for dynamic code loading. This forms
the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and
its variants is passed to \eval, directly or transitively by tainting
their output. This corresponds to \PercentParsedCallSites of the total
\eval call sites and \PercentParsedEvals of the eval calls. We
observed that very few of the eval call (\NbParseFilesRnd in total)
consume the result of calling \c{parse} on a file. Most of the eval
calls consume the result of calling \c{parse} on a string. We also
identified one function in core R, \c{invokeRestartInteractively} that
prompts the user for input, parses it, and passes it to \eval.

>>>>

\mypara{Provenance.}

% Explain tha some generate expression objects and other language objects?
There are various ways to obtain expressions that \eval will evaluate:
\begin{itemize}
    \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
    \item \c{match.call} captures a call and its arguments as expressions.
    \item \c{quote}, \c{enquote} returns its argument unevaluated, as an R expression.
    \item \c{expression}  creates a vector of expression
    objects from text.
    \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
    \item \c{comp} compiles R code into R bytecode.
\end{itemize}

\begin{table}[h]
  \begin{tabular}{|c|c|c|}
      \hline
      & \#sites & \%sites \\
       \hline
         \c{substitute}, \c{quote} & \packageNbOtherLangsSites & \packageOtherLangsSitePercent \\
      \hline
      \c{parse}, \c{str2expression} & \packageNbParseExprsSites & \packageParseExprsSitePercent \\
      \hline
        \c{match.call} &  \packageNbMatchCallExprsSites & \packageMatchCallExprsSitePercent  \\   
      \hline
      \c{expression} & \packageNbNotParseExprsSites & \packageNotParseExprsSitePercent  \\
      \hline
      \c{str2lang} & \packageNbStrlangSites  & \packageStrlangSitePercent  \\
      \hline
      \c{comp} & \packageNbBytecodeSites & \packageBytecodeSitePercent \\
      \hline
  \end{tabular}
\caption{Provenances of \eval expressions. }
\end{table}

% \packageUnknownProvSitePercent : provenance not relavant (values, symbols etc)

\c{eval(parse(...))} can be used for dynamic code loading. This forms
the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and
its variants is passed to \eval, directly or transitively by tainting
their output.  We observed that very few of the eval call (\packageNbParseFromFileSites in total)
consume the result of calling \c{parse} on a file. Most of the eval
calls consume the result of calling \c{parse} on a string. We also
identified one function in core R, \c{invokeRestartInteractively} that
prompts the user for input, parses it, and passes it to \eval.

@@
